### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c17f262cb73c011919904ab00089570c6da12626 🚀

Most previous work has assumed that the presence or richness of key
groups of bat hosts are predictive of coronavirus
diversity.\textsuperscript{2,3} Projecting bat and betacoronavirus
phylogeny over space (fig.~\ref{fig:biogeo}), we find support for the
idea that bat community assembly is directly responsible for a global
mosaic of viral evolution. The distinct groupings (represented by
different colors, symbolizing positions in a subspace formed by the
first two phylogenetic principal components) are essentially equivalent
between the two groups, and can be coarsely delineated as (1) south and
southeast Asia, (2) east Asia (including northern China), west Asia, and
the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
Africa and south America. In some cases, this diverges from expectations
about coronavirus biogeography: for example, previous work has rarely
flagged India as a region of interest, but for both bats and
betacoronaviruses, the subcontinent falls into the same regions as the
southeast Asian peninsula (and indeed, the region is home to known bat
hosts of nobecoviruses, sarbecoviruses, and
merbecoviruses).\textsuperscript{3}

@@

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cospeciation plays a minor role in coronavirus
diversification,\textsuperscript{2} a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,13} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary reservoir of
merbecoviruses;\textsuperscript{3,13} but in the Americas, where
merbecoviruses are the only lineage present, they have only been found
in other bat taxa. At the coarsest scale, these heterogeneities are
lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within broad regions, the component
coevolutionary systems may have very different dynamics.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South
America has the most evolutionary distinct hosts \emph{and} viruses,
followed by secondary hotspots in southeast Asia and the Rift Valley
region have mostly distinct viruses. Some degree of sampling bias may
contribute to these patterns: for example, South-America is one of the
places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{2,14,16} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{17}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) are both the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{17} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,15}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only the subgenus \emph{Merbecovirus} (MERS-like
viruses) has been found in animals in the Americas---an introduction
that is generally presumed to be ancient.\textsuperscript{3} While
comparatively understudied, New World merbecoviruses have been found in
the ghost-faced bats (Mormoopidae), New World leaf-nosed bats
(Phyllostomidae), and free-tailed bats (Molossidae).\textsuperscript{18}
The former two groups are endemic to the Neotropics, while the explosive
adaptive radiations of the latter two (and particularly the
phyllostomids) are responsible for the hotspot of bat diversity in the
Amazon. Together, these clades of New World bats play host to a distinct
regime of betacoronavirus coevolution.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian
continents.\textsuperscript{19} Recent surveillance\textsuperscript{20}
has identified a novel \emph{Betacoronavirus} (in the subgenus
\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
strong proof of principle in model predictions.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward transmission (the probability
that spillovers become epidemics)\textsuperscript{1}. As a proxy for
both, we finally overlaid the risk component from the composite map (see
above) with the proportion of built land, as a proxy for a mix of
habitat disturbance, potential for bat synanthropy or contact with
bridge hosts like livestock,\textsuperscript{21,22} and human population
density and connectivity\textsuperscript{1,23,24}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from the Indochinese peninsula that was poorly
characterized prior to the pandemic.\textsuperscript{25--27} Similarly,
the SARS-CoV outbreak began in Guangdong province in 2002, reaching
humans through small carnivore bridge hosts, but was eventually traced
back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan province;\textsuperscript{28} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{29} MERS-CoV was originally detected in Saudi
Arabia, accompanied by a nearly identical virus sequenced from an
Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{30}, but
is widespread in camels in East Africa and the Middle East, and may have
reached its bridge host decades earlier than originally
supposed;\textsuperscript{31} as a result, the geography of the original
bat-to-camel transmission is still widely regarded as uncertain. All of
these are broadly consistent with the risk factors we identify. Notably,
India and west Africa are additional hotspots that have yet to
experience the emergence of a bat coronavirus into human populations,
but may still be at risk---particularly given known gaps in bat
surveillance,\textsuperscript{17} and a dense population in both regions
with global connectivity. In any of these regions, surveillance on viral
reservoirs can be paired with targeted monitoring of high-risk human
populations (i.e., those with regular wildlife
contact)\textsuperscript{32} for maximum impact.

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{33,34} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species.\textsuperscript{35,36}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to human.\textsuperscript{37} For example,
bats are an essential component of many seed-dispersal
networks.\textsuperscript{38} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{39,40} and
vectors of pathogens that put a risk on human
health.\textsuperscript{41,42} Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses,\textsuperscript{14} bat
coronaviruses,\textsuperscript{2} and even specifically
betacoronaviruses\textsuperscript{15} in both the Amazon and southeast
Asia. While we find that both regions are characterized by highly
divergent host and viral communities, our framework identifies key
differences between the regions. We find that Latin America is a hotspot
of both host and viral distinctiveness, suggesting that this branch of
the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin
America.\textsuperscript{2} In contrast, in southeast Asia, host
richness and viral distinctiveness are high but sharing is low; this
suggests a different type of evolutionary dynamics that could generate
high local diversity of viruses through host switching and viral
recombination (see \emph{e.g.},\textsuperscript{13} as well as the
discovery of recombinant viruses that share genetic material from both
the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{43} Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{15,17} However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region.\textsuperscript{44}

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{45}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{46} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in humans-hosts contacts also increase the risk of emergence of
novel diseases,\textsuperscript{47} so does the changes in landscape
connectivity at local/regional scales.\textsuperscript{48} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{49} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

@@

\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
(2017).}

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
\textbf{17}, 181--192 (2019).}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
Urbanization and disease emergence: Dynamics at the
wildlife--livestock--human interface. \emph{Trends in ecology \&
evolution} \textbf{32}, 55--67 (2017).}

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
330--336 (2022).}

\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}

\leavevmode\hypertarget{ref-Hu2017DisRic}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
bat SARS-related coronaviruses provides new insights into the origin of
SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}

\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
Sinica} \textbf{33}, 104--107 (2018).}

\leavevmode\hypertarget{ref-Memish2013MidEas}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
syndrome coronavirus in bats, Saudi Arabia. \emph{Emerging infectious
diseases} \textbf{19}, 1819 (2013).}

\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
neutralizing antibodies in camels, Eastern Africa, 1983--1997.
\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
1030--1037 (2004).}

>>>>

Most previous work has assumed that the presence or richness of key
groups of bat hosts are predictive of coronavirus
diversity.\textsuperscript{2,3} Projecting bat and betacoronavirus
phylogeny over space (fig.~\ref{fig:biogeo}), we find support for the
idea that bat community assembly is directly responsible for a global
mosaic of viral evolution. The distinct groupings (represented by
different colors, symbolizing positions in a subspace formed by the
first two phylogenetic principal components) are essentially equivalent
between the two groups, and can be coarsely delineated as (1) south and
southeast Asia; (2) east Asia (including northern China), west Asia, and
the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
Africa and Latin America. In some cases, this diverges from expectations
about coronavirus biogeography: for example, previous work has rarely
flagged India as a region of interest, but for both bats and
betacoronaviruses, the subcontinent falls into the same regions as the
southeast Asian peninsula (and indeed, the region is home to known bat
hosts of multiple betacoronavirus subgenera, including nobecoviruses,
sarbecoviruses, and merbecoviruses).\textsuperscript{3}

@@

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cross-species transmission may play a stronger
role in coronavirus diversification than
cospeciation---\textsuperscript{2}a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,13} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary hosts of the
subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,13}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (e.g., Molossidae,
Phyllostomidae) (refs1). At the coarsest scale, these heterogeneities
are lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within broad regions, the component
coevolutionary systems may have very different dynamics.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated (cite2), resulting in a sparser phylogenetic tree, and
artificially inflating distinctiveness; conversely, disproportionate
research effort in eastern China\textsuperscript{16} may have led to a
more complete inventory of the local diversity of coronaviruses, again
inflating these metrics relative to underlying patterns. Even accounting
for these potential biases, though, there is obvious heterogeneity in
betacoronavirus evolutionary distinctiveness that is distinct from
overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{16} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,15}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,17} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae) (refs3). The former two groups and a clade
of the latter are endemic to the Neotropics, while the explosive
adaptive radiations of the phyllostomids are responsible for the hotspot
of bat diversity in the Amazon (refs4). Together, these clades of New
World bats play host to a distinct regime of betacoronavirus
coevolution.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian
continents.\textsuperscript{18} Recent surveillance\textsuperscript{19}
has identified a novel \emph{Betacoronavirus} (in the subgenus
\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
strong proof of principle in model predictions.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward transmission (the probability
that spillovers become epidemics)\textsuperscript{1}. As a proxy for
both, we finally overlaid the risk component from the composite map (see
above) with the proportion of built land, as a proxy for a mix of
habitat disturbance, potential for bat synanthropy or contact with
bridge hosts like livestock,\textsuperscript{20,21} and human population
density and connectivity\textsuperscript{1,22,23}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from the Indochinese peninsula that was poorly
characterized prior to the pandemic.\textsuperscript{24--26} Similarly,
the SARS-CoV outbreak began in Guangdong province in 2002, reaching
humans through small carnivore bridge hosts, but was eventually traced
back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan province;\textsuperscript{27} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{28} MERS-CoV was originally detected in Saudi
Arabia, accompanied by a nearly identical virus sequenced from an
Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{29}, but
is widespread in camels in East Africa and the Middle East, and may have
reached its bridge host decades earlier than originally
supposed;\textsuperscript{30} as a result, the geography of the original
bat-to-camel transmission is still widely regarded as uncertain. All of
these are broadly consistent with the risk factors we identify. Notably,
India and west Africa are additional hotspots that have yet to
experience the emergence of a bat coronavirus into human populations,
but may still be at risk---particularly given known gaps in bat
surveillance,\textsuperscript{16} and a dense population in both regions
with global connectivity. In any of these regions, surveillance on viral
reservoirs can be paired with targeted monitoring of high-risk human
populations (i.e., those with regular wildlife
contact)\textsuperscript{31} for maximum impact.

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{32,33} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species.\textsuperscript{34,35}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to human.\textsuperscript{36} For example,
bats are an essential component of many seed-dispersal
networks.\textsuperscript{37} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{38,39} and
vectors of pathogens that put a risk on human
health.\textsuperscript{40,41} Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses,\textsuperscript{14} bat
coronaviruses,\textsuperscript{2} and even specifically
betacoronaviruses\textsuperscript{15} in both the Amazon and southeast
Asia. While we find that both regions are characterized by highly
divergent host and viral communities, our framework identifies key
differences between the regions. We find that Latin America is a hotspot
of both host and viral distinctiveness, suggesting that this branch of
the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin
America.\textsuperscript{2} In contrast, in southeast Asia, host
richness and viral distinctiveness are high but sharing is low; this
suggests a different type of evolutionary dynamics that could generate
high local diversity of viruses through host switching and viral
recombination (see \emph{e.g.},\textsuperscript{13} as well as the
discovery of recombinant viruses that share genetic material from both
the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{42} Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{15,16} However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region.\textsuperscript{43}

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{44}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{45} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in humans-hosts contacts also increase the risk of emergence of
novel diseases,\textsuperscript{46} so does the changes in landscape
connectivity at local/regional scales.\textsuperscript{47} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{48} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

@@

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
\textbf{17}, 181--192 (2019).}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
Urbanization and disease emergence: Dynamics at the
wildlife--livestock--human interface. \emph{Trends in ecology \&
evolution} \textbf{32}, 55--67 (2017).}

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
330--336 (2022).}

\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}

\leavevmode\hypertarget{ref-Hu2017DisRic}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
bat SARS-related coronaviruses provides new insights into the origin of
SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}

\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
Sinica} \textbf{33}, 104--107 (2018).}

\leavevmode\hypertarget{ref-Memish2013MidEas}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
syndrome coronavirus in bats, Saudi Arabia. \emph{Emerging infectious
diseases} \textbf{19}, 1819 (2013).}

\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
neutralizing antibodies in camels, Eastern Africa, 1983--1997.
\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
1030--1037 (2004).}

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}

\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
virome diversity in bats. \emph{Current Opinion in Virology}
\textbf{52}, 192--202 (2022).}
